---
title: "Create trading days regressors"
output: 
  learnr::tutorial:
    toc_depth: 2
runtime: shiny_prerendered
description: >
  Learn how to create trading days regressors with a user-defined calendar.
---

```{r setup, include=FALSE}
library(learnr)
library(rjd3modelling)
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction 
There are two solutions to create custom working day regressors:

1. Do it from JDemetra+, by creating a personalized calendar then by exporting the regressors. See for example the Jdemetra+ documentation [here](https://jdemetradocumentation.github.io/JDemetra-documentation/pages/case-studies/calendars-main.html) and [there](https://jdemetradocumentation.github.io/JDemetra-documentation/pages/case-studies/calendars-national.html).

2. Create the calendar from R using the package [`rjd3modelling`](https://github.com/palatej/rjd3modelling) (which depends on [`rjd3toolkit`](https://github.com/palatej / rjd3toolkit)). These packages are not on the CRAN, to install them we can for example use the function `remotes::install_github ()`:
```{r, eval=FALSE}
remotes::install_github("palatej/rjd3toolkit",
                 INSTALL_opts = "--no-multiarch")
remotes::install_github("palatej/rjd3modelling",
                 INSTALL_opts = "--no-multiarch")
```

In this tutorial, we will only focus on the second option.
Indeed, the R package is more flexible and allows the creation of less classic regressors than *working days* and *trading days*. 
You can then also use the `rjd3sa` package test for residual trading days.

This this was prepared with the `r packageVersion ('rjd3toolkit')` version of the `rjd3toolkit` and `rjd3modelling`: in new versions, the function names may be different. To install this version you can use the following code:


```{r, include = FALSE}
v_package <- packageVersion('rjd3toolkit')
install_packages <- sprintf("remotes::install_github(\"palatej/%s@v%s\")",
                            c("rjd3toolkit", "rjd3modelling"),
                            v_package)
install_packages <- paste(install_packages, collapse = "\n")
```

```{r, eval=FALSE, code = install_packages}
```

## Creation of a calendar

By default, the trading day regressors of JDemetra+ do not take into account the calendar specificities of the countries: we do not take into account the public holidays.
To do so, you have to create your own calendar where you will assume that one more holiday in the month has the same effect as a Sunday.

A new calendar with the function `calendar.new ()`
```{r}
library(rjd3modelling)
frenchCalendar <- calendar.new()
```
Three functions can be used to add holidays:

1. `calendar.fixedday ()` to add a public holiday that falls on a fixed date. For example, to add May 8:
```{r}
calendar.fixedday(frenchCalendar, month =  5,
                  day = 8)
```

2. `calendar.easter ()` to add a holiday whose day depends on Easter: the `offset` parameter allows you to specify the number of days before (if negative) or after Easter (if positive). For example, to add Pentecost which takes place 59 days after Easter:

```{r}
calendar.easter(frenchCalendar,
                offset = 59)
```

3. `calendar.holiday ()` which allows you to add public holidays in relation to dates already known in JDemetra+ (see table below). As for the `calendar.easter()` function, the `offset` parameter allows you to specify the position of the desired day in relation to the pre-specified holiday (by default` offset = 0`, the holiday coincides with the day pre-specified). For example, to add the new year:
```{r}
calendar.holiday(frenchCalendar, "NEWYEAR")
```


```{r, prespecifieholidays, echo=FALSE}
library(kableExtra)
holidays <- data.frame(Event = c("NEWYEAR", "SHROVEMONDAY", "SHROVETUESDAY", 
"ASHWEDNESDAY", "MAUNDYTHURSDAY", "GOODFRIDAY", "EASTER", "EASTERMONDAY", 
"ASCENSION", "PENTECOST", "WHITMONDAY", "CORPUSCHRISTI", "JULIANEASTER", 
"MAYDAY", "ASSUMPTION", "HALLOWEEN", "ALLSAINTDAY", "ARMISTICE", 
"CHRISTMAS"), Definition = c("Fixed holiday, falls on January, 1.", "Moving holiday, falls on Monday before Ash Wednesday (48 days before Easter Sunday).", 
"Moving holiday, falls on Tuesday before Ash Wednesday (47 days before Easter Sunday).", 
"Moving holiday, occurring 46 days before Easter Sunday.", "Moving holiday, falls on the Thursday before Easter.", 
"Moving holiday, falls on the Friday before Easter.", "Moving holiday, varies between March, 22 and April, 25.", 
"Moving holiday, falls on the day after Easter.", "Moving holiday, celebrated on Thursday, 39 days after Easter.", 
"Moving holiday, celebrated 49 days after Easter Sunday.", "Moving holiday, falling on the day after Pentecost.", 
"Moving holiday, celebrated 60 days after Easter Sunday.", "", 
"Fixed holiday, falls on May, 1.", "Fixed holiday, falls on August, 15.", 
"Fixed holiday, falls on October, 31.", "Fixed holiday, falls on November, 1.", 
"Fixed holiday, falls on November, 11.", "Fixed holiday, falls on December, 25."
))
kbl(holidays,booktabs = TRUE,caption = "Pre-specified holidays",row.names = FALSE) %>%
  kable_paper("hover", full_width = F)
```


Now your turn! Create a calendar that contains all the public holidays of France:
```{r frenchcalendar, exercise = TRUE}
# see for example: https://demarchesadministratives.fr/actualites/calendrier-des-jours-feries-2019-2020-2021
frenchCalendar <- calendar.new()
```

```{r frenchcalendar-solution}
frenchCalendar <- calendar.new()
calendar.holiday(frenchCalendar, "NEWYEAR")
calendar.holiday(frenchCalendar, "EASTERMONDAY")
calendar.holiday(frenchCalendar, "MAYDAY")
calendar.fixedday(frenchCalendar, 5, 8)
calendar.holiday(frenchCalendar, "WHITMONDAY")
calendar.fixedday(frenchCalendar, 7, 14)
calendar.holiday(frenchCalendar, "ASSUMPTION")
calendar.holiday(frenchCalendar, "ALLSAINTDAY")
calendar.holiday(frenchCalendar, "ARMISTICE")
```


## Create trading day regressors

The general working day correction model can be written as follows:
$$
X_t = \sum_{i=1}^{7} \alpha_i N_{it} + \varepsilon_t
$$
With:

- $N_{it}$ the number of days of Mondays ($i = 1$), ..., Sundays and holidays ($i = 7$)

- $\alpha_i$ the effect of a day of type $i$

To avoid the problems of multi-collinearity, one rewrites the model by using a modality of reference (here Sunday). The regressors are also seasonally adjusted by removing the long-term average:
$$X_t =  \sum_{i=1}^{6} \beta_i (N_{it} - N_{7t}) + 
\bar{\alpha} \underbrace{(N_t - \bar{N}_t)}_{LY_t} +  \varepsilon_t$$
This model can be simplified if by making assumptions about the effects of working days:

- The *working days* hypothesis corresponds to the case where we assume that all the days of the week (Monday to Friday) have the same effect ($\alpha_1 = \dots = \alpha_5$), Saturdays and Sundays (and holidays) have the same effect ($\alpha_6 = \alpha_7$) and are used as a contrast variable.

- The *trading days* hypothesis corresponds to the case where it is assumed that all days have a different effect and Sundays (and holidays) are used as a constrate variable.

Under Jdemetra+ we can only use these two hypotheses but `rjd3modelling` allows to build other types of regressors with user-defined groups.


More generally, when using a contrast variable, the regressors $TDR_{t, i}$ associated to the group $i$ are calculated as follows:
$$
TDR_{t,i} = \underbrace{\sum_{j\in\text{group }i}N_{jt}}_{
\text{nb of days of group }i
} - 
\frac{\sum_{j\in\text{groupe }i}1}{\sum_{j\in\text{group }0}1}
\times
\underbrace{\sum_{j\in\text{group }0}N_{jt}}_{
\text{nb of days in contrast group}
}
$$
They are also corrected from long-term mean but we don't go into details here.


In the case of *working days*, there are 2 days in the contrast group (Saturday and Sunday, $\sum_{j\in\text{group }0}1=2$) and 5 days in group 1 (Monday until Friday, $\sum_{j\in\text{groupe }1}1=5$). 
At month $t$, the trading day regressors are therefore equal to the number of weekdays in the month, minus $5/2\times$  the number of weekend days.

Trading days regressors can be created from 2 functions: `htd()`, to create them from a specific calendar; and `td()`, default parameters.
In these functions, the most important parameter is `groups` that allows you to make assumptions about the days. 
It is a vector of length 7 (the number of days of the week), each element indicates to which group the associated day of the week corresponds. 
The contrast variable is associated with group 0.
For example, `groups = c (1,2,3,4,5,6,0)` corresponds to *trading days* and `groups = c(1,1,1,1,1,0,0)` corresponds to *working days*.

For example:
```{r}
groups <- c(1, 2, 3, 4, 5, 6, 0)
frequency <- 12
start <- c(2000,1)
wkd <- htd(frenchCalendar, frequency = frequency, start = start, length = 12*35,
         groups = groups)
wkd <- ts(wkd, start = start, frequency = frequency)
```


Compare the  *working days* regressors created with the French calendar to the one without assumption on public holidays (`td()` function).

```{r regressors, exercise = TRUE}
frenchCalendar <- calendar.new()
```

```{r regressors-hint}
# define Saturday and Sunday as contrast
groups <- c(1, 1, 1, 1, 1, 0, 0)
# Compare with td() function
wkd_def <- td(frequency = frequency, start = start, length = 12*35,
         groups = groups)
```

```{r regressors-solution}
groups <- c(1, 1, 1, 1, 1, 0, 0)
frequency <- 12
start <- c(2000,1)
wkd <- htd(frenchCalendar, frequency = frequency, start = start, length = 12*35,
         groups = groups)
wkd <- ts(wkd, start = start, frequency = frequency)
wkd_def <- td(frequency = frequency, start = start, length = 12*35,
         groups = groups)
wkd_def <- ts(wkd_def, start = start, frequency = frequency)
data <- ts.union(wkd, wkd_def)
plot(data, col = c("orange","black"),
     plot.type = "single")
```

## Easter variables

```{r easter-find, echo=FALSE}
question("Use the function `easter.dates()` to find the date of Easter in 1935",
  answer("April 12", message = "Nope, April 12 was Easter in 1936!"),
  answer("March 31"),
  answer("April 21", correct = TRUE),
  answer("April 15")
)
```

